### 2018-03-24 morus.tex
...
Confirmed correlation for two halves of 1280 trail
added part of Yu's observations to tex

\usepackage{geometry}
\usepackage{amsmath,amssymb}
\usepackage{ifthen}
\usepackage{subcaption}

@@

\noindent
2 approximations for $S^{2,2}_{0}$:
\begin{align*}
  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus
  C^4_{19} \tag{corr. $2^{-9}$} \\
\intertext{Combined ciphertext-only approximation:}
  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-16.5}$}
\end{align*}

@@

\begin{figure}
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
    \printstate
  \end{tikzpicture}%}}}
  \caption{\cipher{MiniMORUS} state update function (1 step = 5 rounds).}
  \label{fig:minimorus}
\end{figure}

@@

  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, delta]
      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
      (lll4) -- (W44) node[below] {$i+b_4$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\delta_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, epsil]
      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
      (lll2) -- (W42) node[below] {$i+b_2$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\varepsilon_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
    \small
    \begin{align*}
      \intertext{\cipher{MiniMORUS-640}:}
      \alpha_i:      && C_i &\to S^{0}_{i+5} \\
      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+31} \\
      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+13} \\
      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+7}
      \intertext{\cipher{MiniMORUS-1280}:}
      \alpha_i:      && C_i &\to S^{0}_{i+13} \\
      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+46} \\
      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+4} \\
      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+38}
    \end{align*}
    \caption*{\cipher{MiniMORUS} instances}
  \end{subfigure}
  \caption{\cipher{MiniMORUS} linear approximation fragments.}
  \label{fig:minimorus}
\end{figure}

@@

      \draw (\S{4}{2}) node[above] {\normalsize weight 7 (not 9)};
    \end{scope}

@@

      \draw (\S{5}{2}) node[above] {\normalsize weight 9 (not 11)};
      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
    \end{scope}

@@

      \draw (\S{4}{2}) node[above] {\normalsize weight 7 (not 9)};
      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
    \end{scope}

@@

      \draw (\S{5}{2}) node[above] {\normalsize weight 9 (not 11)};
    \end{scope}
  \end{tikzpicture}
  \caption{\cipher{MiniMORUS}: Two approximations for $S^{2,2}_0$}
\end{figure}

\begin{figure}

  \caption{Suspicious things I'd like to verify with full \cipher{MORUS}}
\end{figure}

>>>>

\usepackage{geometry}
\usepackage{amsmath,amssymb}
\usepackage{ifthen}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[section]{placeins}
\usepackage[hidelinks]{hyperref}

@@

@@

\section{Description of \cipher{MORUS} and \cipher{MiniMORUS}}

\begin{figure}[h]
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{tikzpicture}[xscale=1.0,yscale=1.5]%{{{
    \printstate
  \end{tikzpicture}%}}}
  \caption{\cipher{MiniMORUS} state update function (1 step = 5 rounds).}
  \label{fig:minimorus}
\end{figure}

\clearpage

\section{Linear Approximations for \cipher{MiniMORUS}}

Basic idea using the fragments in \autoref{fig:minimorusfragments} and a fixed message $M=0$:
\begin{itemize}
    \item $\alpha_i$ approximates (one bit of) state word $S^0$ with the ciphertext
    \item $\beta_i$ approximates $S^1$ using $S^0$ and the ciphertext
    \item $\gamma_i$ approximates $S^4$ using two approximations of $S^1$ in consecutive steps
    \item $\delta_i$ approximates $S^2$ using two approximations of $S^4$ in consecutive steps
    \item $\varepsilon_i$ approximates $S^0$ using two approximations of $S^2$ in consecutive steps
\end{itemize}
Thus, we obtain a ciphertext-only approximation by appropriately connecting
$1 \times \alpha$ with $1 \times \varepsilon$, $2 \times \delta$, $4 \times \gamma$, $8 \times \beta$, $8 \times \alpha$.
The attack is essentially independent of the rotation constants $b_i$.
The resulting biased ciphertext expression is listed for both
\cipher{MiniMORUS-640} and \cipher{MiniMORUS-1280} in \autoref{tab:miniapproximations}.
Since each fragment has weight 1, we would expect a total weight of 24;
however, the approximations are not independent (\autoref{fig:todo}).
The actual weight is 16, and experiments confirm an even higher correlation of $2^{-15.5}$.

%\noindent
%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-640}:
%\begin{align*}
%  S^{2,2}_0 &= C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12} \tag{corr. $2^{-7}$} \\
%  S^{2,2}_0 &= C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19} \tag{corr. $2^{-9}$} \\
%\intertext{Combined ciphertext-only approximation:}
%  0 &= C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus
%  C^3_{6,12,14,20} \oplus C^4_{19} \tag{corr. $\approx 2^{-15.5}$}
%\end{align*}
%\bigskip
%
%\noindent
%2 approximations for $S^{2,2}_{0}$ of \cipher{MiniMORUS-1280}:
%\begin{align*}
%  S^{2,2}_0 &= C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50} \tag{corr. $2^{-7}$} \\
%  S^{2,2}_0 &= C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24} \tag{corr. $2^{-9}$} \\
%\intertext{Combined ciphertext-only approximation:}
%  0 &= C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24} \tag{corr. $\approx 2^{-??}$}
%\end{align*}

\begin{table}[hb!]
  \caption{Verification of approximations}
  \label{tab:miniapproximations}
  \begin{tabular}{@{}llSSS@{}}
    \toprule
    & & \multicolumn{3}{@{}c@{}}{Weight} \\
    \cmidrule{3-5}
    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS-640}}
             & {Exp.} & {Bool.} & {Meas.} \\
    \midrule
    $\chi_1$ & $S^{2,2}_0 = C^0_{27} \oplus C^1_{0, 8, 26} \oplus C^2_{7,13,31} \oplus C^3_{12}$
             & 7    & 7     & 7 \\
    $\chi_2$ & $S^{2,2}_0 = C^1_{2} \oplus C^2_{1,7,15,27} \oplus C^3_{6,14,20} \oplus C^4_{19}$
             & 9    & 9     & 9 \\
    $\chi$   & $0 = C^0_{27} \oplus C^1_{0, 2, 26, 8} \oplus C^2_{1,13,15,27,31} \oplus C^3_{6,12,14,20} \oplus C^4_{19}$
             & 16   & 16    & 15.5 \\
    \midrule
    \multicolumn{2}{@{}l}{Approximations for \cipher{MiniMORUS-1280}} \\
    \midrule
    $\chi_1$ & $S^{2,2}_0 = C^0_{51} \oplus C^1_{0, 33, 55} \oplus C^2_{4,37,46} \oplus C^3_{50}$
             & 7    & 7     & 7 \\
    $\chi_2$ & $S^{2,2}_0 = C^1_{25} \oplus C^2_{7,29,38,51} \oplus C^3_{11,20,42} \oplus C^4_{24}$
             & 9    & 9     & 9 \\
    $\chi$   & $0 = C^0_{51} \oplus C^1_{0, 25, 33, 55} \oplus C^2_{4,7,29,37,38,46,51} \oplus C^3_{11,20,42,50} \oplus C^4_{24}$
             & 16   & 16    & {??} \\
    \bottomrule
  \end{tabular}
\end{table}

@@

  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, delta]
      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
      (lll4) -- (W44) node[below] {$i+b_4$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\delta_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, epsil]
      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
      (lll2) -- (W42) node[below] {$i+b_2$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\varepsilon_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
    \small
    \begin{align*}
      \intertext{\cipher{MiniMORUS-640}:}
      \alpha_i:      && C_i &\to S^{0}_{i+5} \\
      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+31} \\
      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+13} \\
      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+7}
      \intertext{\cipher{MiniMORUS-1280}:}
      \alpha_i:      && C_i &\to S^{0}_{i+13} \\
      \beta_i:       && C_i, S^0_i, S^1_i &\to 0 \\
      \gamma_i:      && S^1_i, S^4_i &\to S^1_{i+46} \\
      \delta_i:      && S^4_i &\to S^2_i, S^4_{i+4} \\
      \varepsilon_i: && S^2_i &\to S^0_i, S^2_{i+38}
    \end{align*}
    \caption*{\cipher{MiniMORUS} instances}
  \end{subfigure}
  \caption{\cipher{MiniMORUS} linear approximation fragments.}
  \label{fig:minimorusfragments}
\end{figure}

\begin{figure}
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, beta]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (tlll-1) -- (W-20) node[above] {$i$}
      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
      %(xnd-1) -- (and-1)
      (W40) node[below] {\phantom{$i$}}
      ;
    \draw[trail, gamma]
      %(W-21) node[above] {$i$}
      (txor-1) -- (xnd1)
      (xnd1) -- (xor1) (xor1) -- (lll1)
      %(xnd1) -- (and1)
      (xor1) -- (txor1)
      (txor1) -- (W-24) node[above] {$i$}
      (lll1) -- (W41) node[below] {$i+b_1$}
      ;
    %\draw[trail, beta,  dotted] (and-1) -- (and-1-|and1);
    %\draw[trail, gamma, dotted] (and1)  -- (and-1-|and1) node[above, black] {=};
    \draw[trail, beta]  (xnd-1) -- ++(-.5,0) coordinate (tmp) -- (tmp|-xnd1);
    \draw[trail, gamma] (tmp|-xnd1) node[below, black] {=} -- (xnd1);
  \end{tikzpicture}%}}}
  \caption*{\cipher{MiniMORUS}: weight 0 (not 2)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
  \end{subfigure}
  \qquad
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, beta]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (tlll-1) -- (W-20) node[above] {$i$}
      (xor-1) -- (txor-1) %(txor-1) -- (W-21) node[above] {$i$}
      (xnd-1) -- (and-1)
      (W40) node[below] {\phantom{$i$}}
      ;
    \draw[trail, gamma]
      %(W-21) node[above] {$i$}
      (txor-1) -- (xnd1)
      (xnd1) -- (xor1) (xor1) -- (lll1)
      (xnd1) -- (and1)
      (xor1) -- (txor1)
      (txor1) -- (W-24) node[above] {$i$}
      (lll1) -- (W41) node[below] {$i+b_1$}
      ;
    \draw[trail, beta,  dashed] (and-1.east|-tanB-1) -- (tanB-1);
    \draw[trail, gamma, dashed] (tanB-1)  -- (tanB1) -- (tanB1-|and1.east);
  \end{tikzpicture}%}}}
  \caption*{\cipher{MORUS}: weight $4 \times 1$ (not $4 \times 2$)} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
  \end{subfigure}
  \caption{Weight of $\beta_i \oplus \gamma_i$ for \cipher{MiniMORUS} and \cipher{MORUS}.}
  \label{fig:minimorusfragments}
\end{figure}

TODO: check $\alpha_i \oplus \beta_i \oplus \gamma_i$ in step 2 of $\cipher{MORUS-640}$
(maybe not relevant due to overall complexity\dots)

@@

      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
    \end{scope}

@@

      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
      \draw (\S{0}{2}) node[below] {\normalsize \cipher{MiniMORUS-640}};
    \end{scope}

@@

      \draw (\S{4}{2}) node[above] {\normalsize $\chi_1$: weight 7 (not 11)};
      \draw (\S{5}{2}) node[above] {\normalsize \cipher{MiniMORUS-1280}};
    \end{scope}

@@

      \draw (\S{5}{2}) node[above] {\normalsize $\chi_2$: weight 9 (not 13)};
    \end{scope}
  \end{tikzpicture}
  \caption{\cipher{MiniMORUS}: Two approximations for $S^{2,2}_0$}
\end{figure}

\clearpage

\section{Linear Approximations for \cipher{MORUS}}

\section{Differential Attack on the Finalization}

\section{Key-Recovery Attack in Nonce-Misuse Setting}
